Sketch Recognition and Understanding

### 2009-08-10 trunk/draft/System.tex

\section{Classification}%%Feature Set
\label{sec:Classification}
The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}. SVM is a binary classifier that maximizes the margin between the decision boundary and the support vectors. To enable SVM to learn multi classes we used one versus one strategy for combining set of binary classifier. This strategy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes. The final result is based on a number of votes each shape has from all classifiers. We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy.

\subsection{Support Vector Machines}

The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{} in 1995. It is gainig increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a sperating plane between two different classes which are represented using a set of n-dimentional vectors. The generated plane is used as a good classifier to the unseen exmaples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{}. To calcuate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generenate the maximum margin possilbe.

@@

The margin depends on the Support vecotrs that are bumpt into the planes.  This is a primal problem Quadratic Programming (QP) problem. The concept of duality can also upsed to wields the problem to a  Largrnigian dual QP problem in Equation \ref{eq:dual}. 
% maximumm margin between parallel 
\begin{equation}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
  \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
\label{eq:dual}
\end{equation}
Both the dual form and the primal form weilds to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).

@@

\subsubsection{Linearly in seperable case}
 So far we have assumed that the two datasets are linearly separable. Th this is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail. As illustrated in Figure \ref{}, if the points are not linearly seperable the primal supporitng plane will fails. Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains. Consider the linear inseperable problem in Figure\ref{}, Ideally we would like no points to be missclassified and no points to fall in the margin. But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane. Any point fallihng on the wrong side of tis supprting plane is considered to be an error. We want to maximize the margin and minimize the error.

 This could be accomplished through minor change in the supporging plane QP problem. A nonegative slack or error variable $z_i$ is added to each constraint and then added as weighted penalty term in the objective as follows:
 \begin{equation}
\min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
s.t.  \quad y_i(\omega \cdot x_i -b)+z_i \geq 1  \\
    z_i \geq 0  i=1,\cdots,m
\label{eq:inseperable}
\end{equation}

@@

\begin{equation}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
\label{eq:dual}
\end{equation}

\subsubsection {Non linear function Using Kernels}
%The foundations of Support Vector Machines (SVM) have
\subsubsection{Summary of SVM}
The resulting SVM method can be summarized as follows 
\begin{enumerate}
	\item Select paramter C representign the tradeoff between minimizign the training set error and maximizing the margin. Select the kernel function and any kernel paramters. For example for the radial basis function kernel, one must select width of the gaussian $\gamma$. 
	\item Solve dual QP Equation \ref{} or an alterantive SVM formulation using an approriate quadratic programming or linear programming alogrithm. 
	\item Recover th primal threshold variable $b$ using the support vectors. 
	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
\end{enumerate}

The paramters in step 1. are slected using cross valiation if sufficent data are avaliable. %However, recent model  selectin strategies can

\subsubsection{ SVM in Machine Learning }

\susubsection { SVM vs. Other Learning Methods}

>>>>

\section{Classification}%%Feature Set
\label{sec:Classification}
The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}. SVM is a binary classifier that maximizes the margin between the decision boundary and the support vectors. To enable SVM to learn multi classes we used one versus one strategy for combining set of binary classifier. This strategy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes. The final result is based on a number of votes each shape has from all classifiers. We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy. Section \ref{Sec:SVMdetail} discribe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{} in 1995. It is gainig increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a sperating plane between two different classes which are represented using a set of n-dimentional vectors. The generated plane is used as a good classifier to the unseen exmaples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{}. To calcuate the margin, two different hyperplanes are constructed, one plane on each side of the data, these two planes are then moved to generenate the maximum margin possilbe.

@@

The margin depends on the Support vecotrs that are bumpt into the planes.  This is a primal problem Quadratic Programming (QP) problem. The concept of duality can also upsed to wields the problem to a  Largrnigian dual QP problem in Equation \ref{eq:dual}. 
% maximumm margin between parallel 
\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
  \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
  \end{array} 
\label{eq:dual}
\end{equation}
Both the dual form and the primal form weilds to the same plane $\omega=\sum_{i=1}^{m}y_i\alpha_i x_i$ and threshold $b$ determined by the support vectors (for which $\alpha_i\geq 0$).

@@

\subsubsection{Linearly Inseperable Case}
 So far we have assumed that the two datasets are linearly separable. Th this is not true, the startegy of constructing the plane that bisect the two closest points of the convex hulls will fail. As illustrated in Figure \ref{}, if the points are not linearly seperable the primal supporitng plane will fails. Since the QP task is not feasible for ethe linearly inseperable case , then we need to relax the constrains. Consider the linear inseperable problem in Figure\ref{}, Ideally we would like no points to be missclassified and no points to fall in the margin. But we must relax the constrain that insure that each point in on the approiate side of tis supporitng plane. Any point fallihng on the wrong side of tis supprting plane is considered to be an error. We want to maximize the margin and minimize the error.

 This could be accomplished through minor change in the supporging plane QP problem. A nonegative slack or error variable $z_i$ is added to each constraint and then added as weighted penalty term in the objective as follows:
 \begin{equation}
 \begin{array}{c}
\min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
s.t.  \quad y_i(\omega \cdot x_i -b)+z_i \geq 1  \\
    z_i \geq 0  i=1,\cdots,m
\end{array} 
\label{eq:inseperable}
\end{equation}

@@

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_jx_ix_j}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dual}
\end{equation}

\subsubsection {Non linear function Using Kernels}
If we consider a classification problem that cannot be seperated using a linear space (See Figure \ref{} ). A quadratic or  cubic function will work better in some casses. A classific method of converting a linear classification algoirhtm inot a nonolinear classification algoorithm is to simply add addional attribute to the data that are nonlinear functions of the original data. Existing linar classification algorithms can be then applied to the expanded dataset feature space producing non linear functions in the original input space. To construct a quadratic disceriminant in a two dimensional vector space with attributes $r$ and $s$, simply map the original two dimensional input space $\left[r,s\right]$ to the five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminiant in that space. specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then 
\begin{equation}
\begin{array}{l}
	x=[r,s] \\
	\omega \cdot x=\omega_1 r+ \omega_2 s \\
	\downarrow \\
	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\

	 \omega \cdot \vartheta(x)= \omega_1 r+ \omega_2 s + \omega_3 rs+ \omega_4 r^2+\omega_5 r^2 
\end{array}
\label{eq:kernel}
\end{equation}  
The resulting classification function, 
\begin{equation}
\begin{array}{l}
f(x)=sign\left( \omega \cdot\vartheta(x)-b \right) \\
= sign\left( \omega_1 r+ \omega_2 s + \omega_3 rs+ \omega_4 r^2+\omega_5 r^2 -b \right) 
\end{array}
\label{eq:kern2}
\end{equation}  
is linear in the mapped five dimensional feature space but is quadratic in the two dimensional input space.

For higher dimentional datasets, this nonlinear mapping method has two potential problems stemming from the fact taht dimensionality of the feature space explones exponentially. The first problem is that overfitting becomes a problem. SVM are largely immune to this porlblem since they rely on margin maximization. Provided an appropriate value of paaramter $C$ is chosen. The second concern is that it is not practical to actually compute $\vartheta(x)$. SVMs get around this issue through teh use of kernels.

Exampine what happens when athe nonlinear mapping is introduced to equation \ref{eq:dual}. Let us define : $\vartheta(x):R^2\rightarrow R^5 \quad n>> n$  we need to optimize

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically rigorous magie known as hilbern schmidt kernels, first applied to SVMs in \cite{}.  By Mercer's theorem, we knoew that for ccertain mapping $\theta$ and any two points $u$ and $v$, the inner product of the the mapped points can be evalutated using the kernel function without ever explicity knowing ht emapping , e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernel are being developed to fit domian specific requriments. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
		Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\

	\end{tabular}
\end{center}
To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective insted of the original dot product. Thus by changing kernels we can get different hightly nonlinear classifiers. No algorithmic changes are required from linear clase other than substitue of a kernel evaluation for simple dot product. All the benefits fo the original linear SVM methods are maintained. We can train a highly nonlinear classification function such as a ploynomial or a raadial basis function machine, or a sigmoidal neural network using robust, efficient algorithsm that have no problem with local minima. By using kernel substitution a linear algorithm (only capable of handling seperable data) can be turned into a general nonlinear algorithm. 
%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

%The foundations of Support Vector Machines (SVM) have
\subsubsection{Classification}
The resulting SVM method can be summarized as follows 
\begin{enumerate}
	\item Select paramter C representign the tradeoff between minimizign the training set error and maximizing the margin. Select the kernel function and any kernel paramters. For example for the radial basis function kernel, one must select width of the gaussian $\gamma$. 
	\item Solve dual QP Equation \ref{} or an alterantive SVM formulation using an approriate quadratic programming or linear programming alogrithm. 
	\item Recover th primal threshold variable $b$ using the support vectors. 
	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
\end{enumerate}

The paramters in step 1. are selected using cross valiation if sufficent data are avaliable. %However, recent model  selectin strategies can

\subsubsection{ SVM in Machine Learning }
SVM has been successfully applied to a number of applications ranging from particle identification \cite{}. Face detection  \cite{} and text categorization \cite{} to engine knock detection 
\subsubsection {SVM vs. Other Learning Methods}

\subsubsection {Multi Class Classifiers}
\subsubsection {SVM Implementations}
Typically an SVM approach requires the solution of a QP or LP problem. LP and QP type problems have been extensively studied inthe field of mathematical programming. These can be divided into three categories techniques in which kernel compoinents are evaluated and discarded during learning. Decomposition methods in which an evolving subset of data is used and new optimization approaches that specifically exploit the structure of the SVM problem.

For the first category the most obvioius  approach is to sequentially update the $\alpha_i$  and this is the approach used by the kernel  Adtron (KA) algorithm \cite{}.

Chunking and decomposition methods optimize the SVM with respect to subsets. Rather than sequentially updateing the $\alpha_i$ the alternative is to upodate the $\alpha_i$ in parallel but using only a subset or wokring set of data at each stage. There are many successful code based on these decomposition strategies. SVM codes available online such as SVMTorch \cite{} and SVMLight \cite{} use these working set strategies.

The third approach is to directly attack the SVM  problem from and optimization prespective and create algorithms that explicity exploits the structure of teh problem. Keerthi et al. \cite{} proposed a very effective algorithm based on the dual geometry of finding the two closest points in the convex hulls.

\subsubsection {Summary Of SVM}
